Below-ground sink limitation alters growth and carbon balance of Eucalyptus seedlings

### 2015-12-07 manuscript/manuscript_pve.Rmd
fixed chapters figure #s
tables now automated

```{r, echo=FALSE}
allomfig <- "2.1"
airvarsfig <- "2.2"
allocfig <- "2.3"
asatfig <- "2.4"
achemfig <- "2.5"
massmodelfig <- "2.6"
massmodelfig2 <- "2.S1"
```

@@

\
(2)
$$R = \Sigma(R_c~M_c)$$
where R~c~ is published or local tissue respiration rates of fine roots, coarse roots or stems (Table S1, g C g mass^-1^ d^-1^) and M~c~ is the standing biomass of each component (g). R~leaf~ is represented in the calculation of C~day~ (described below). The change in individual component biomass (M~c~), here solved on a daily time step, is given by

@@

\
It was further necessary to calculate a self-shading parameter ($\sigma$~s~) when scaling leaf *A* with total plant leaf area. This was accomplished by utilizing 61 previously digitized Eucalyptus seedlings, covering 5 total species which include *E. tereticornis*, from Duursma et al. [-@duursma2012light] to run in 'YplantQMC' package [@YplantQMC] in R to build a 3D plant structure based on digitized metrics of plant allometry and crown structure. Inputting the same treatment specific physiological parameters listed above, 'YplantQMC' outputs total *A*, using total leaf area, for seedlings assuming self-shading as well as for a full sun large horizontal leaf.  The ratio of total *A* with self-shading to horizontal leaf was then used to calculate $\sigma$~s~ for each of the 61 digitized seedlings, independently for each of the seven soil volume treatments. Next, the linear relationship between $\sigma$~s~ and total leaf area was for determined across digitized seedlings, within each treatment. For the growth model, $\sigma$~s~ was then predicted for each daily time step using the previous days cumulative leaf area and this value  was then applied to C~day,i~. All default parameters used in model simulations are reported in Table S1.

@@

\
Total seedling biomass at harvest was significantly different across container volumes (P < 0.001) and between container treatments and free seedlings (P < 0.001, Table 1). On average, harvested biomass of free seedlings was 84% higher than seedlings in containers. Plant biomass was positively correlated with total leaf area across all treatments (R^2^ = 0.97, P < 0.001). Differences in biomass partitioning to leaves, stems and roots were not different across soil volumes after variation in seedling biomass across treatments was factored in the analysis (Figure `r allocfig`a,b). Across all treatments, the final harvest root:shoot biomass ratio was conserved in these seedlings which exhibited a slightly higher shoot than root mass ($\overline{x}$ = 0.904, 95% CI = [0.846,1.119]) and a near identical ratio of leaf to fine root mass (Figure `r allocfig`c).

\
**Table 1**.Responses of plant and leaf characteristics of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean(standard error) for each treatment. Seedling mass and leaf $\delta$^13^C values are from final harvest. Values of leaf starch, sugars, nitrogen and SLA represent overall means across measurement campaigns (n=6). Different letters represent significant differences between treatments. The volume effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings.

@@

\
Overall, SRL was higher in seedlings in containers compared to free seedlings but only in some of the container size treatments (Table 2, P = 0.009). Fine root length density was significantly higher in the two smallest container sizes and was the lowest in the largest container size (Table 2, P < 0.001). Over the duration of the experiment SLA~f~ was higher in free seedlings, but was not different across containers sizes (Table 1, P < 0.001) and this pattern was evident beginning in the first gas exchange measurement campaign (P < 0.001).

\
**Table 2**. Responses of root characteristics of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean(standard error) for each treatment. All values are from the final harvest. Values for FRLD were only calculated for seedlings in containers as free seedlings had potentially unlimited soil volume to exploit. Different letters represent significant differences between treatments. The volume effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings, except for FRLD which represents only differences between seedlings in containers.
```{r table2, echo=FALSE, warning=FALSE}
dust(table2)%>%
  sprinkle_colnames("Volume (L)", 
                        "Root Nitrogen (%)", 
                        "SRL (m g^-1^)",
                        "FRLD (m dm^-3^)") %>%
  sprinkle(rows= 1, part="head",bold=TRUE)%>%
  sprinkle_print_method("markdown")
```

## Leaf and root chemistry
Leaf N~f~ was significantly higher in free seedlings and the largest container volume compared to the smaller container volumes at the onset of gas exchange measurements (6th week, P < 0.001). Throughout the remainder of the experiment the smallest container volume had a significant reduction in leaf N~f~ compared to other soil volumes, while free seedlings maintained the highest leaf N~f~ (Table 1, P < 0.001). Leaf starch content in the smallest container was ca. double  that of free seedlings (P = 0.039), while leaf soluble sugars did not differ across treatments throughout the experiment (Table 1). Differences in leaf starch between the free seedling and the smallest container were evident during the first gas exchange campaign (P = 0.001). Root N was higher in free seedlings compared to seedlings in containers but only for some of the container size treatments (Table 2).

## Gas exchange and photosynthetic parameters
At the first measurement campaign, both A~sat~ and A~max~ were significantly higher in the free seedling treatment compared to seedlings in containers (both P < 0.001). Across all measurement campaigns mean A~sat~ (Figure `r asatfig`) and A~max~ (Table 3) were consistently higher in free seedlings than in containers (26 % and 29 %, respectively). The relationships between photosynthetic capacity, leaf starch and leaf N on a mass basis was marginally significant (P = 0.058) but A~max~ on a mass basis was highly correlated to both leaf N content and leaf starch (both P < 0.001). We used predictions from the linear mixed effect model equation to visualize these relationship of A~max~ to either leaf N content or leaf starch at multiple bin levels (n=5) of the co-variate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments A~max~ was higher when leaf N was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). A~max~  was also lower when leaf starch was high as higher leaf N often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, A~max~ was positively correlated with final harvest biomass across all seedlings (P < 0.001).

@@

\
Both J~max~ and Vc~max~ were significantly higher in free seedlings (30 % and 26 %, respectively) than container-grown seedlings with little variation between container volume treatments (Table 3). Leaf dark respiration rates were not significantly different across soil volumes (Table 3). The g~1~ parameter, generated for each seedling from the Medlyn et al [-@medlyn2011reconciling] optimal stomatal conductance model, was lowest in the free seedling treatment and was marginally different across soil volume treatments (Table 3).

\
Neither $\Psi$~pd~ nor $\Psi$~l~ were different across treatments, with mean values of -0.27 and -1.2 MPa across all seedlings, respectively. Although g~s~ in free seedlings was generally higher than those in containers (Table 3. P < 0.001), the mean rates for all seedlings were high at 0.37 mol m^-2^ s^-1^ and did not change throughout the course of the experiment. Additionally, leaf $\delta$^13^C at final harvest was not different across treatments (Table 1). Combined these indices provide strong evidence that water stress was not apparent on these seedlings throughout the experiment. Soil N at harvest was not different across soil volumes ($\bar{x}$ = 4.5 %), with minimal decreases from pre-planting value ($\bar{x}$ = 4.9 %). This indicates that nutrient leaching from free seedlings or from draining of containers following natural rainfall events did not differ between treatments.

>>>>

```{r, echo=FALSE}
allomfig <- "2.1"
airvarsfig <- "2.2"
allocfig <- "2.3"
asatfig <- "2.4"
achemfig <- "2.5"
massmodelfig <- "2.6"
massmodelfig2 <- "2.S1"

pvetab1 <- "2.1"
pvetab2 <- "2.2"
pvetab3 <- "2.3"
pvetabS <- "2.S3"

@@

\
(2)
$$R = \Sigma(R_c~M_c)$$
where R~c~ is published or local tissue respiration rates of fine roots, coarse roots or stems (Table `r pvetabS`, g C g mass^-1^ d^-1^) and M~c~ is the standing biomass of each component (g). R~leaf~ is represented in the calculation of C~day~ (described below). The change in individual component biomass (M~c~), here solved on a daily time step, is given by

@@

\
It was further necessary to calculate a self-shading parameter ($\sigma$~s~) when scaling leaf *A* with total plant leaf area. This was accomplished by utilizing 61 previously digitized Eucalyptus seedlings, covering 5 total species which include *E. tereticornis*, from Duursma et al. [-@duursma2012light] to run in 'YplantQMC' package [@YplantQMC] in R to build a 3D plant structure based on digitized metrics of plant allometry and crown structure. Inputting the same treatment specific physiological parameters listed above, 'YplantQMC' outputs total *A*, using total leaf area, for seedlings assuming self-shading as well as for a full sun large horizontal leaf.  The ratio of total *A* with self-shading to horizontal leaf was then used to calculate $\sigma$~s~ for each of the 61 digitized seedlings, independently for each of the seven soil volume treatments. Next, the linear relationship between $\sigma$~s~ and total leaf area was for determined across digitized seedlings, within each treatment. For the growth model, $\sigma$~s~ was then predicted for each daily time step using the previous days cumulative leaf area and this value  was then applied to C~day,i~. All default parameters used in model simulations are reported in Table `r pvetabS`.

@@

\
Total seedling biomass at harvest was significantly different across container volumes (P < 0.001) and between container treatments and free seedlings (P < 0.001, Table `r pvetab1`). On average, harvested biomass of free seedlings was 84% higher than seedlings in containers. Plant biomass was positively correlated with total leaf area across all treatments (R^2^ = 0.97, P < 0.001). Differences in biomass partitioning to leaves, stems and roots were not different across soil volumes after variation in seedling biomass across treatments was factored in the analysis (Figure `r allocfig`a,b). Across all treatments, the final harvest root:shoot biomass ratio was conserved in these seedlings which exhibited a slightly higher shoot than root mass ($\overline{x}$ = 0.904, 95% CI = [0.846,1.119]) and a near identical ratio of leaf to fine root mass (Figure `r allocfig`c).

\
**Table 2.1**.Responses of plant and leaf characteristics of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean(standard error) for each treatment. Seedling mass and leaf $\delta$^13^C values are from final harvest. Values of leaf starch, sugars, nitrogen and SLA represent overall means across measurement campaigns (n=6). Different letters represent significant differences between treatments. The volume effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings.

@@

\
Overall, SRL was higher in seedlings in containers compared to free seedlings but only in some of the container size treatments (Table `r pvetab2`, P = 0.009). Fine root length density was significantly higher in the two smallest container sizes and was the lowest in the largest container size (Table `r pvetab2`, P < 0.001). Over the duration of the experiment SLA~f~ was higher in free seedlings, but was not different across containers sizes (Table `r pvetab1`, P < 0.001) and this pattern was evident beginning in the first gas exchange measurement campaign (P < 0.001).

\
**Table 2.2**. Responses of root characteristics of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean(standard error) for each treatment. All values are from the final harvest. Values for FRLD were only calculated for seedlings in containers as free seedlings had potentially unlimited soil volume to exploit. Different letters represent significant differences between treatments. The volume effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings, except for FRLD which represents only differences between seedlings in containers.
```{r table2, echo=FALSE, warning=FALSE}
dust(table2)%>%
  sprinkle_colnames("Volume (L)", 
                        "Root Nitrogen (%)", 
                        "SRL (m g^-1^)",
                        "FRLD (m dm^-3^)") %>%
  sprinkle(rows= 1, part="head",bold=TRUE)%>%
  sprinkle_print_method("markdown")
```

## Leaf and root chemistry
Leaf N~f~ was significantly higher in free seedlings and the largest container volume compared to the smaller container volumes at the onset of gas exchange measurements (6th week, P < 0.001). Throughout the remainder of the experiment the smallest container volume had a significant reduction in leaf N~f~ compared to other soil volumes, while free seedlings maintained the highest leaf N~f~ (Table `r pvetab1`, P < 0.001). Leaf starch content in the smallest container was ca. double  that of free seedlings (P = 0.039), while leaf soluble sugars did not differ across treatments throughout the experiment (Table `r pvetab1`). Differences in leaf starch between the free seedling and the smallest container were evident during the first gas exchange campaign (P = 0.001). Root N was higher in free seedlings compared to seedlings in containers but only for some of the container size treatments (Table `r pvetab2`).

## Gas exchange and photosynthetic parameters
At the first measurement campaign, both A~sat~ and A~max~ were significantly higher in the free seedling treatment compared to seedlings in containers (both P < 0.001). Across all measurement campaigns mean A~sat~ (Figure `r asatfig`) and A~max~ (Table `r pvetab3`) were consistently higher in free seedlings than in containers (26 % and 29 %, respectively). The relationships between photosynthetic capacity, leaf starch and leaf N on a mass basis was marginally significant (P = 0.058) but A~max~ on a mass basis was highly correlated to both leaf N content and leaf starch (both P < 0.001). We used predictions from the linear mixed effect model equation to visualize these relationship of A~max~ to either leaf N content or leaf starch at multiple bin levels (n=5) of the co-variate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments A~max~ was higher when leaf N was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). A~max~  was also lower when leaf starch was high as higher leaf N often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, A~max~ was positively correlated with final harvest biomass across all seedlings (P < 0.001).

@@

\
Both J~max~ and Vc~max~ were significantly higher in free seedlings (30 % and 26 %, respectively) than container-grown seedlings with little variation between container volume treatments (Table `r pvetab2`). Leaf dark respiration rates were not significantly different across soil volumes (Table `r pvetab2`). The g~1~ parameter, generated for each seedling from the Medlyn et al [-@medlyn2011reconciling] optimal stomatal conductance model, was lowest in the free seedling treatment and was marginally different across soil volume treatments (Table `r pvetab2`).

\
Neither $\Psi$~pd~ nor $\Psi$~l~ were different across treatments, with mean values of -0.27 and -1.2 MPa across all seedlings, respectively. Although g~s~ in free seedlings was generally higher than those in containers (Table `r pvetab2`, P < 0.001), the mean rates for all seedlings were high at 0.37 mol m^-2^ s^-1^ and did not change throughout the course of the experiment. Additionally, leaf $\delta$^13^C at final harvest was not different across treatments (Table `r pvetab1`). Combined these indices provide strong evidence that water stress was not apparent on these seedlings throughout the experiment. Soil N at harvest was not different across soil volumes ($\bar{x}$ = 4.5 %), with minimal decreases from pre-planting value ($\bar{x}$ = 4.9 %). This indicates that nutrient leaching from free seedlings or from draining of containers following natural rainfall events did not differ between treatments.
